---
title: "Homework1"
author: "Kirk Vanacore"
date: "9/8/2021"
output: html_document
---

## Bayesian Computation in Stan and R 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load/Setup STRAN

```{r cars}
library(rstan) # observe startup messages
options(mc.cores = parallel::detectCores()) # Use all local cores (does this work for all r packages?)
rstan_options(auto_write = TRUE) # saves stan to hard drive
Sys.setenv(TZ = "America/Toronto") # this deals with a time zone bug issue that was causes an error when running the model

```

# Example 1: Eight Schools
This is an example in Section 5.5 of Gelman et al (2003), which studied coaching effects from eight schools. For simplicity, we call this example "eight schools."

## Write STAN file
We start by writing a Stan program for the model in a text file. If you are using RStudio version 1.2.x or greater, click on File -> New File -> Stan File . Otherwise, open your favorite text editor. Either way, paste in the following and save your work to a file called schools.stan in R's working directory (which can be seen by executing getwd())
You can also embed plots, for example:

```
// saved as schools.stan
data {
  int<lower=0> J;         // number of schools 
  real y[J];              // estimated treatment effects
  real<lower=0> sigma[J]; // standard error of effect estimates 
}
parameters {
  real mu;                // population treatment effect
  real<lower=0> tau;      // standard deviation in treatment effects
  vector[J] eta;          // unscaled deviation from mu by school
}
transformed parameters {
  vector[J] theta = mu + tau * eta;        // school treatment effects
}
model {
  target += normal_lpdf(eta | 0, 1);       // prior log-density
  target += normal_lpdf(y | theta, sigma); // log-likelihood
}

```
Be sure that your Stan programs ends in a blank line without any characters including spaces and comments.

In this Stan program, we let theta be a transformation of mu, eta, and tau instead of declaring theta in the parameters block, which allows the sampler will run more efficiently . 

### Prepare data
We can prepare the data (which typically is a named list) in R with:
```{r, echo=FALSE}
rm(list = ls())
schools_dat <- list(J = 8, 
                    y = c(28,  8, -3,  7, -1,  1, 18, 12),
                    sigma = c(15, 10, 16, 11,  9, 11, 10, 18))
```

And we can get a fit with the following R command. Note that the argument to file = should point to where the file is on your file system unless you have put it in the working directory of R in which case the below will work.


```{r}
fit <- stan(file = 'schools.stan', data = schools_dat)
```

##unexpected WARNINGS: 
Warning message:
In utils::install.packages("openssl", repos = "https://cran.rstudio.com/") :
installation of package ‘openssl’ had non-zero exit status


## Output 
```{R}
print(fit)

```

# Plot Paramteres
```{R}

plot(fit)

```


```{R}

pairs(fit, pars = c("mu", "tau", "lp__"))

```


```{R}

la <- extract(fit, permuted = TRUE) # return a list of arrays 
mu <- la$mu 

### return an array of three dimensions: iterations, chains, parameters 
a <- extract(fit, permuted = FALSE) 

### use S3 functions on stanfit objects
a2 <- as.array(fit)
m <- as.matrix(fit)
d <- as.data.frame(fit)

```