---
title: "Homework 2"
author: "Kirk Vanacore"
date: "9/8/2021"
output: html_document
---
## Bayesian Computation in Stan and R 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Load/Setup STRAN

```{r}
library(rstan) # observe startup messages
options(mc.cores = parallel::detectCores()) # Use all local cores (does this work for all r packages?)
rstan_options(auto_write = TRUE) # saves stan to hard drive
Sys.setenv(TZ = "America/Toronto") # this deals with a time zone bug issue that was causes an error when running the model

```

This is an example in Section 5.5 of Gelman et al (2003), which studied coaching effects from eight schools. For simplicity, we call this example "eight schools."

## Write STAN file
We start by writing a Stan program for the model in a text file. If you are using RStudio version 1.2.x or greater, click on File -> New File -> Stan File . Otherwise, open your favorite text editor. Either way, paste in the following and save your work to a file called schools.stan in R's working directory (which can be seen by executing getwd())
You can also embed plots, for example:

```
// saved as schools.stan
data {
  int<lower=0> J;         // number of schools 
  real y[J];              // estimated treatment effects
  real<lower=0> sigma[J]; // standard error of effect estimates 
}
parameters {
  real mu;                // population treatment effect
  real<lower=0> tau;      // standard deviation in treatment effects
  vector[J] eta;          // unscaled deviation from mu by school
}
transformed parameters {
  vector[J] theta = mu + tau * eta;        // school treatment effects
}
model {
  target += normal_lpdf(eta | 0, 1);       // prior log-density
  target += normal_lpdf(y | theta, sigma); // log-likelihood
}
```

Be sure that your Stan programs ends in a blank line without any characters including spaces and comments.

In this Stan program, we let theta be a transformation of mu, eta, and tau instead of declaring theta in the parameters block, which allows the sampler will run more efficiently . 
 


### Prepare data
We can prepare the data (which typically is a named list) in R with:
```{r}
rm(list = ls())
schools_dat <- list(J = 9, 
                    y = c(5.41,  5, 6.29,  10.95, 9.53, 7.28, 6.59, 2.72, 6.94), 
                    sigma = c(0.33, 0.25, 0.2, 0.12,  0.19, 0.23, 0.24, 0.22, 0.17)) # treatment effects for each school

condition_dat <- list(J = 4, 
                    y = c(7.41, 6.01, 7.21, 7.33), 
                    sigma = c(0.18, 0.18, 0.13, 0.18)) # treatment effects for each school

#WHY DO WE USE THE SE? SHOULDN"T SIGMA BE THE SD (its it because the SE takes into account sample size?)

```

And we can get a fit with the following R command. Note that the argument to file = should point to where the file is on your file system unless you have put it in the working directory of R in which case the below will work.

# School Models
RESEARCH QUESTIONS:
Does the postest vary between schools?
```{r}
fit <- stan(file = 'schools.stan', data = schools_dat)
```



## Output 
```{R}
print(fit)
```

## Plot Paramteres
```{R}

plot(fit, pars = c("mu",
                   "tau",
                   "eta[1]", # differences between the overall mean and the cluster mean 
                   "eta[2]",
                   "eta[3]",
                   "eta[4]",
                   "eta[5]",
                   "eta[6]",
                   "eta[7]",
                   "eta[8]",
                   "eta[9]"
                   ))


```

# 
yi ~ N(theta1i, sigmai)
theta ~ N(mu, tau)



```{R}

pairs(fit, pars = c("mu", # means 
                    "tau", # between cluster variance
                    "lp__" #?
                    ))

```

# Condition Models
RESEARCH QUESTIONS:
Does the posttest vary between Conditions?
```{r}
fit2 <- stan(file = 'schools.stan', data = condition_dat)
```


## Plot Paramteres
```{R}

plot(fit2, pars = c("mu",
                   "tau",
                   "eta[1]",
                   "eta[2]",
                   "eta[3]",
                   "eta[4]"
                   ))


```
```{R}
pairs(fit, pars = c("mu", # mean
                    "tau", # between cluster variance
                    "lp__")) # some version of the posteior 
```


# other code
```{R}

la <- extract(fit, permuted = TRUE) # return a list of arrays 
mu <- la$mu 

### return an array of three dimensions: iterations, chains, parameters 
a <- extract(fit, permuted = FALSE) 

### use S3 functions on stanfit objects
a2 <- as.array(fit)
m <- as.matrix(fit)
d <- as.data.frame(fit)

```