---
title: "Homework 3"
author: "Kirk Vanacore"
date: "9/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Load/Setup STRAN

```{r}
library(rstan) # observe startup messages
options(mc.cores = parallel::detectCores()) # Use all local cores (does this work for all r packages?)
rstan_options(auto_write = TRUE) # saves stan to hard drive
Sys.setenv(TZ = "America/Toronto") # this deals with a time zone bug issue that was causes an error when running the model
library(dplyr)
```


# Homework 3

1) make a stan_data object like this:
> stan_data= list(N=nrow(dat), y=dat$post.math.total,pretest=dat$pre.math.total)

(modify as necessary--the point is that you are creating a list of all of the data elements that you put in the data{} part of the stan model)

2) fit the model like this (again, modify as necessary)
> fit=stan("regression1.stan", data=stan_data)

3) check out the results 

extra credit: put in another predictor

###Save in file .stan
```
// The input data is a vector 'y' of length 'N'.
data {
  int<lower=0> N;
  vector[N] y;
  vector[N] pretest_MC;
}

// yi = b0 + b1xi1 + ei
// ei ~ N(0, sigma)
// yi ~ N(b0 + b1xi1, sigma) <- This is the modle


// The parameters accepted by the model. Our model
// accepts two parameters 'mu' and 'sigma'.
parameters {
  real<lower=0> sigma;
  real b0;
  real b1;
}

// The model to be estimated. We model the output
// 'y' to be normally distributed with mean 'mu'
// and standard deviation 'sigma'.
model {
  y ~ normal(b0 + b1*pretest_MC, sigma);
}

```


## Prepare data
```{r}

assess <-read.csv("Assessment_merged_2021_07_16_state_assessment_N=4321 - Sheet1.csv", na.strings = c(""))
colnames(assess)
# Clean data
assess <- assess %>%
  filter(is.na(post.percentage_math_score) == F,
         is.na(pre.percentage_math_score) == F,
         
    # remove students who are not in original random assignment
    is.na(rdm_condition) == F,
    # remove students who are in resource
     assess$rdm_condition != "Dragon-Resource" & assess$rdm_condition != "FH2T-Resource",
    # S03 drop schools 
     final_school_id != "S03" & initial_school_id != "S03"
    & final_school_id != "S07" & initial_school_id != "S07"

  )


# 
hist(assess$post.percentage_math_score)

assess$pre.percentage_math_score_MC <- assess$pre.percentage_math_score- mean(assess$pre.percentage_math_score)


stan_data1 <-  list(N=nrow(assess), 
                   y=assess$post.percentage_math_score,
                   pretest_MC=assess$pre.percentage_math_score_MC)


```

And we can get a fit with the following R command. Note that the argument to file = should point to where the file is on your file system unless you have put it in the working directory of R in which case the below will work.

# Model1: Predicting post test from prestest
```{r}

fit1 <- stan(file = 'HW3stanScript.stan',
            data = stan_data1)

```



## Output 
```{R}
print(fit1)
```


# Model 2 

## Stan Code
```
//
// This Stan program defines a simple model, with a
// vector of values 'y' modeled as normally distributed
// with mean 'mu' and standard deviation 'sigma'.
//
// Learn more about model development with Stan at:
//
//    http://mc-stan.org/users/interfaces/rstan.html
//    https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started
//

// The input data is a vector 'y' of length 'N'.
data {
  int<lower=0> N;
  vector[N] y;
  vector[N] pretest_MC;
  vector[N] FH2T;
  vector[N] Dragon;
  vector[N] ASSISTments;

  
}

// The parameters accepted by the model. Our model
// accepts two parameters 'mu' and 'sigma'.
parameters {
  real<lower=0> sigma;
  real b0;
  real b1;
  real b2;
  real b3;
  real b4; 
}

// The model to be estimated. We model the output
model {
  y ~ normal(b0 
    + b1*pretest_MC 
    + b2*FH2T 
    + b3*Dragon 
    + b4*ASSISTments , sigma);
}

```

## Prepare Data
```{R}

# Dummy Code
table(assess$rdm_condition)
assess$FH2T <- ifelse(assess$rdm_condition == "FH2T", 1, 0)
table(assess$rdm_condition, assess$FH2T)

assess$ASSISTments <- ifelse(assess$rdm_condition == "ASSISTments", 1, 0)
table(assess$rdm_condition, assess$ASSISTments)


assess$Dragon <- ifelse(assess$rdm_condition == "Dragon", 1, 0)
table(assess$rdm_condition, assess$Dragon)
nrow(assess)


stan_data2 <-  list(N=nrow(assess), 
                   y=assess$post.percentage_math_score,
                   pretest_MC=assess$pre.percentage_math_score_MC,
                   FH2T = assess$FH2T,
                   ASSISTments = assess$ASSISTments,
                   Dragon = assess$Dragon)

```


# Model2: Estimaiting treatment effects
```{r}

fit2 <- stan(file = 'HW3stanScript2.stan',
            data = stan_data2)

```

```{R}
print(fit2)
```

How do I evaluate the Coefficients For example: b2 CIs (maybe that is not what they are) overlap zero. Does this mean that we can't say that it is 95% probable that the effect of FH2T is not zero? I'm I think about this incorrectly because are not doing null hypotheses testing.

```{R}
plot(fit2)
```