---
title: "Homework 4"
author: "Kirk Vanacore"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data
```{r}
data<-read.csv("data_assignment5.csv")
```
## Models - One Exposure (one row per student)
### Random Intercept Only

$$\hat{y}\sim\log(\beta_{0j}+\beta_{2} Treatement_{1i}+\beta_2 PriorAccuracy_{2j}+e_ij)$$
  $$\beta_0 = \gamma_{00} + \gamma_{02} ProblemDifficulty_{j} + \mu_{0i} $$
_I'm not sure that the Problem Difficulty is in the right place._

#### glmer code
```{r}
# m1 <- glmer(
#   next_problem_correctness_complete ~
#       treatment +
#         prior_accuracy + 
#         problem_avg_accuracy +
#   (1|problem_id) ,
#   data = data,
#   family = binomial)
# summary(m1)
```
_Note that problem_avg_accuracy is a measure of problem difficulty._


#### STAN CODE

``` 
data {
  
}

parameters {
  
}

transformed parameters {
  
}


model {
 
}
```


### Random Intercept and Random Effect Treatment

$$\hat{y}\sim\log(\beta_{0j}+\beta_{2} Treatement_{1i}+\beta_2 PriorAccuracy_{2j}+e_ij)$$ 
  $$\beta_0 = \gamma_{00} + \gamma_{02} ProblemDifficulty_{j} + \mu_{0i} $$
  $$\beta_1 = \gamma_{10} + \mu_{1i} $$

#### glmer code
```{r}
# m1 <- glmer(
#   next_problem_correctness_complete ~
#       treatment +
#         prior_accuracy + 
#         problem_avg_accuracy +
#   (1+treatment|problem_id) ,
#   data = data,
#   family = binomial)
# summary(m1)
```
_Note that problem_avg_accuracy is a measure of problem difficulty._


#### STAN CODE

``` 
data {
  
}

parameters {
  
}

transformed parameters {
  
}


model {
 
}
```
### Random Intercept and Random Effect Treatment with cross Level Interaction

$$\hat{y}\sim\log(\beta_{0j}+\beta_{2} Treatement_{1i}+\beta_2 PriorAccuracy_{2j}+e_ij)$$ 
  $$\beta_0 = \gamma_{00} + \gamma_{02} ProblemDifficulty_{j} + \mu_{0i} $$
  $$\beta_1 = \gamma_{10} + \gamma_{12} ProblemDifficulty_{j} + \mu_{1i} $$
  
#### glmer code
```{r}
# m1 <- glmer(
#   next_problem_correctness_complete ~
#       treatment +
#         prior_accuracy + 
#         problem_avg_accuracy +
#         problem_avg_accuracy*treatment +
#   (1+treatment|problem_id) ,
#   data = data,
#   family = binomial)
# summary(m1)
```
_Note that problem_avg_accuracy is a measure of problem difficulty._


#### STAN CODE

``` 
data {
  
}

parameters {
  
}

transformed parameters {
  
}


model {
 
}
```
